### 2019-11-26 ic-journal/equa.tex
It is recommended that "cf." be used only to suggest a comparison, and the word "see" be used to point to a source of information.
https://en.wikipedia.org/wiki/List_of_Latin_abbreviations

In the CCS syntax, 
a recursion $\recu A  P$ acts as a binder for $A$ in the body $P$. 
This gives rise, in the expected manner, to the notions of 
\emph{free} and \emph{bound} recursion variables in a CCS expression. 
For instance,  $X$ is free in $a.X + \recu Y (b.Y)$ while $Y$
is bound; \hl{And} in $a.X + \recu X (b.X)$, $X$ is both free and bound.
A \hl{term} without free variables is \hl{called} a \emph{process}.
% This setting does not cause any ambiguity, but sometimes leads to more
% complex proofs.} (c.f. Section~\ref{sec:multivariate} for more details.)

In this paper (and the formalisation work), we use the agent
variables also as \emph{equation variables}. This eliminates the need of
another type for  equations, and we can reuse the existing
variable substitution operation (c.f.\ the   SOS rule for the Recursion in
Fig.~\ref{f:LTSCCS}) for the substitution of equation variables.
For example, the result of substituting variable $X$ with $\nil$ in $a.X +
\recu X (b.X)$,  written $(a.X + \recu X (b.X)) \sub {\nil} X$, is
$a.\nil + \recu X (b.X)$ (with the part $\recu X (b.X)$
untouched). \Multivariate substitutions are written in the same syntax,
e.g. $E \sub {\til P} {\til X}$. Whenever $\til X$ is clear from the
context, we may also write $E[\til P]$ instead of $E \sub {\til P} {\til
  X}$ (and $E[P]$ for $E \sub P X$ if there is a single equation
variable $X$).

>>>>

In the CCS syntax, 
a recursion $\recu A  P$ acts as a binder for $A$ in the body $P$. 
This gives rise, in the expected manner, to the notions of 
\emph{free} and \emph{bound} recursion variables in a CCS expression. 
For instance,  $X$ is free in $a.X + \recu Y (b.Y)$ while $Y$
is bound; \hl{And} in $a.X + \recu X (b.X)$, $X$ is both free and bound.
A \hl{term} without free variables is \hl{called} a \emph{process}.
% This setting does not cause any ambiguity, but sometimes leads to more
% complex proofs.} (see Section~\ref{sec:multivariate} for more details.)

In this paper (and the formalisation work), we use the agent
variables also as \emph{equation variables}. This eliminates the need of
another type for  equations, and we can reuse the existing
variable substitution operation (cf.~the SOS rule for the Recursion in
Fig.~\ref{f:LTSCCS}) for the substitution of equation variables.
For example, the result of substituting variable $X$ with $\nil$ in $a.X +
\recu X (b.X)$,  written $(a.X + \recu X (b.X)) \sub {\nil} X$, is
$a.\nil + \recu X (b.X)$ (with the part $\recu X (b.X)$
untouched). \Multivariate substitutions are written in the same syntax,
e.g. $E \sub {\til P} {\til X}$. Whenever $\til X$ is clear from the
context, we may also write $E[\til P]$ instead of $E \sub {\til P} {\til
  X}$ (and $E[P]$ for $E \sub P X$ if there is a single equation
variable $X$).

### 2019-11-26 ic-journal/multivariate.tex
It is recommended that "cf." be used only to suggest a comparison, and the word "see" be used to point to a source of information.
https://en.wikipedia.org/wiki/List_of_Latin_abbreviations

Proposition~\ref{prop:transFV} \hl{is essential in the proofs of the
  multivariable versions of all unique-solution
  theorems.
In comparison with} the above proof, the proof of the similar
result for bound variables is \hl{much} easier:
\begin{proposition}
\label{prop:transBV}
if $E \arr{\mu} E'$ then $\bv{E'} \subseteq \bv{E}$, or formally:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{u} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{BV} \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLBoundVar{E}\hfill{[TRANS_BV]}
\end{alltt}
\end{proposition}
\begin{proof}
Following the proof of Prop.~\ref{prop:transFV}, we do action
induction on $E \arr{\mu} E'$ and simplify all subgoals by set-theoretic
facts. The only subgoal left is:
\begin{alltt}
        \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{E}}
   ------------------------------------
    0.  \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}}
\end{alltt}
Note the difference with the proof of Prop.~\ref{prop:transFV}: now
we have \texttt{INSERT} in place of \texttt{DELETE}. Then, we directly
use the assumption and the transitivity of $\subseteq$ and try to
prove the following goal instead:
\begin{alltt}
        \HOLinline{\HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{E}}
   ------------------------------------
    0.  \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}}
\end{alltt}
This is an easy goal, due to the following lemma which is
 proved by induction on $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{E} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLConst{BV} \ensuremath{(}\ensuremath{[}\ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLBoundVar{E} \HOLSymConst{\HOLTokenUnion{}} \HOLConst{BV} \ensuremath{\HOLBoundVar{E}\sp{\prime}}\hfill{[BV_SUBSET]}
\end{alltt}
Taking $E ' = \recu X E$ in the above lemma, we have
    \HOLinline{\HOLConst{BV}\\\;\ensuremath{(}\ensuremath{[}\HOLConst{rec}\\\;\HOLFreeVar{X}\\\;\HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\HOLFreeVar{E}\ensuremath{)}\\\;\HOLSymConst{\HOLTokenSubset{}}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}\\\;\HOLSymConst{\HOLTokenUnion{}}\\\;\HOLConst{BV}\\\;\ensuremath{(}\HOLConst{rec}\\\;\HOLFreeVar{X}\\\;\HOLFreeVar{E}\ensuremath{)}},
    the right side equals to \HOLinline{\HOLConst{BV}\\\;\HOLFreeVar{E}\\\;\HOLSymConst{\HOLTokenUnion{}}\\\;\ensuremath{(}\HOLFreeVar{X}\\\;\HOLConst{INSERT}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}\ensuremath{)}} $=$
    \HOLinline{\HOLFreeVar{X}\\\;\HOLConst{INSERT}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}}. Thus
the    proof can 
 be completed.
\end{proof}

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLTokenDefEquality{}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Note that the variable substitution only occurs on free
variables. In the case of $\recu X E$: if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must continue the
substitution on $E$ using a reduced map~--- without $X$~--- so that all
occurences of $X$ in $E$ are correctly bypassed.
Since \HOLinline{\HOLConst{CCS_SUBST}} only runs
once on each subterm, the possible free variables in  \HOLinline{\HOLFreeVar{Ps}} are never
 substituted. 
% Thus the process is indeed
% order-independent. 
Below we present some key lemmas about
\HOLinline{\HOLConst{CCS_SUBST}}, omitting the proofs:\footnote{Hereafter,
  some new logical constants from HOL's \texttt{pred_set}
and \texttt{list} theories are used (c.f.~\cite{holdesc} for more
details.): \HOLinline{\HOLConst{DISJOINT}} denotes set disjointness;
``\HOLinline{\HOLConst{set}\\\;\HOLFreeVar{Xs}}'' is the set converted from the list \HOLinline{\HOLFreeVar{Xs}};
\HOLinline{\HOLConst{ALL_DISTINCT}} says
all elements of a list are distinct with each other; \HOLinline{\HOLConst{MAP}} is
the mapping function from one list to another; \HOLinline{\HOLConst{EVERY}} means
each element of a list satisfies a predicate; \HOLinline{\HOLConst{ZIP}} creates a
new list from two lists of the same length, and each element of the new list is a pair of
elements, each from one of the old lists.}

@@

Recall \hl{that} in Section~\ref{ss:context}, \univariate contexts and their
guarded companions are defined as predicates over $\lambda$-expressions
of type $CCS\rightarrow CCS$. These predicates, such as
\HOLinline{\HOLConst{CONTEXT}}
(contexts/expressions), \HOLinline{\HOLConst{WG}} (weak guarded contexts),
\HOLinline{\HOLConst{SG}} (guarded contexts) and \HOLinline{\HOLConst{SEQ}} (sequential
contexts), are all defined inductively, i.e.~they are built from some ``holes''
in a bottom-up way. For instance, \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{\mid}}\\\;\HOLFreeVar{P}\ensuremath{)}}
holds because \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}}
holds as a base case of the inductive definition of \HOLinline{\HOLConst{WG}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{a}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}\hfill{[WG1]}
\end{alltt}
For any agent variable $X$, we have by $\beta$-reduction:
\HOLinline{\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}\ensuremath{)}\\\;\ensuremath{(}\HOLConst{var}\\\;\HOLFreeVar{X}\ensuremath{)}\\\;\HOLSymConst{\ensuremath{=}}\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}} (or ``$l.X + P$'' in textbook
notation) is weakly guarded (c.f.~Def.~\ref{def:guardness}).

@@

Not every expression fit with CCS syntax is a context: any free
variable in concern must not occur inside any recusion operator
(c.f. the footnote at the beginning of Section~\ref{ss:SysEq}). In the
extreme case, if the set of free variables of an expression, say $E$,
is disjoint with a list of variables $\til X$, then this expression is
for sure a context:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{E}\hfill{[disjoint_imp_context]}
\end{alltt}
On the other hand, for any context (w.r.t.~$\til X$) of the form $\recu Y E$, we can
conclude that, the set of free variables of $E$ excluding $Y$, is
disjoint with $\til X$, i.e.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{context} \HOLFreeVar{Xs} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{Y} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLFreeVar{E} \HOLConst{DELETE} \HOLFreeVar{Y}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)}\hfill{[context_rec]}
\end{alltt}
Notice that, the above result is the best possible one. We cannot
conclude \HOLinline{\HOLConst{DISJOINT}\\\;\ensuremath{(}\HOLConst{FV}\\\;\HOLFreeVar{E}\ensuremath{)}\\\;\ensuremath{(}\HOLConst{set}\\\;\HOLFreeVar{Xs}\ensuremath{)}}, because $Y$ as a bound
variable of $\recu Y E$ may be also a free variable in $\til X$. We
also cannot conclude \HOLinline{\HOLConst{context}\\\;\HOLFreeVar{Xs}\\\;\HOLFreeVar{E}}, because in $\recu Y E$
it is possible that the bound variable $Y$ occurs inside another
recursion operator, then in $E$ it now becomes free. (See $\recu Y
(\recu Z (a.Y + b.Z) + c.Y)$ for such an example.)

>>>>

Proposition~\ref{prop:transFV} \hl{is essential in the proofs of the
  multivariable versions of all unique-solution
  theorems.
In comparison with} the above proof, the proof of the similar
result for bound variables is \hl{much} easier:
\begin{proposition}
\label{prop:transBV}
if $E \arr{\mu} E'$ then $\bv{E'} \subseteq \bv{E}$, or formally:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{u} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{BV} \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLBoundVar{E}\hfill{[TRANS_BV]}
\end{alltt}
\end{proposition}
\begin{proof}
Following the proof of Prop.~\ref{prop:transFV}, we do action
induction on $E \arr{\mu} E'$ and simplify all subgoals by set-theoretic
facts. The only subgoal left is:
\begin{alltt}
        \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{E}}
   ------------------------------------
    0.  \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}}
\end{alltt}
Note the difference with the proof of Prop.~\ref{prop:transFV}: now
we have \texttt{INSERT} in place of \texttt{DELETE}. Then, we directly
use the assumption and the transitivity of $\subseteq$ and try to
prove the following goal instead:
\begin{alltt}
        \HOLinline{\HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{E}}
   ------------------------------------
    0.  \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}}
\end{alltt}
This is an easy goal, due to the following lemma which is
 proved by \hl{doing} induction on $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{E} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLConst{BV} \ensuremath{(}\ensuremath{[}\ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLBoundVar{E} \HOLSymConst{\HOLTokenUnion{}} \HOLConst{BV} \ensuremath{\HOLBoundVar{E}\sp{\prime}}\hfill{[BV_SUBSET]}
\end{alltt}
Taking $E ' = \recu X E$ in the above lemma, we have
    \HOLinline{\HOLConst{BV}\\\;\ensuremath{(}\ensuremath{[}\HOLConst{rec}\\\;\HOLFreeVar{X}\\\;\HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\HOLFreeVar{E}\ensuremath{)}\\\;\HOLSymConst{\HOLTokenSubset{}}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}\\\;\HOLSymConst{\HOLTokenUnion{}}\\\;\HOLConst{BV}\\\;\ensuremath{(}\HOLConst{rec}\\\;\HOLFreeVar{X}\\\;\HOLFreeVar{E}\ensuremath{)}},
    the right side equals to \HOLinline{\HOLConst{BV}\\\;\HOLFreeVar{E}\\\;\HOLSymConst{\HOLTokenUnion{}}\\\;\ensuremath{(}\HOLFreeVar{X}\\\;\HOLConst{INSERT}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}\ensuremath{)}} $=$
    \HOLinline{\HOLFreeVar{X}\\\;\HOLConst{INSERT}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}}. Thus
the    proof can 
 be completed.
\end{proof}

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLTokenDefEquality{}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Note that the variable substitution only occurs on free
variables. In the case of $\recu X E$: if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must continue the
substitution on $E$ using a reduced map~--- without $X$~--- so that all
occurences of $X$ in $E$ are correctly bypassed.
Since \HOLinline{\HOLConst{CCS_SUBST}} only runs
once on each subterm, the possible free variables in  \HOLinline{\HOLFreeVar{Ps}} are never
 substituted. 
% Thus the process is indeed
% order-independent. 
Below we present some key lemmas about
\HOLinline{\HOLConst{CCS_SUBST}}, omitting the proofs:\footnote{Hereafter,
  some new logical constants from HOL's \texttt{pred_set}
and \texttt{list} theories are used (see~\cite{holdesc} for more
details.): \HOLinline{\HOLConst{DISJOINT}} denotes set disjointness;
``\HOLinline{\HOLConst{set}\\\;\HOLFreeVar{Xs}}'' is the set converted from the list \HOLinline{\HOLFreeVar{Xs}};
\HOLinline{\HOLConst{ALL_DISTINCT}} says
all elements of a list are distinct with each other; \HOLinline{\HOLConst{MAP}} is
the mapping function from one list to another; \HOLinline{\HOLConst{EVERY}} means
each element of a list satisfies a predicate; \HOLinline{\HOLConst{ZIP}} creates a
new list from two lists of the same length, and each element of the new list is a pair of
elements, each from one of the old lists.}

@@

Recall \hl{that} in Section~\ref{ss:context}, \univariate contexts and their
guarded companions are defined as predicates over $\lambda$-expressions
of type $CCS\rightarrow CCS$. These predicates, such as
\HOLinline{\HOLConst{CONTEXT}}
(contexts/expressions), \HOLinline{\HOLConst{WG}} (weak guarded contexts),
\HOLinline{\HOLConst{SG}} (guarded contexts) and \HOLinline{\HOLConst{SEQ}} (sequential
contexts), are all defined inductively, i.e.~they are built from some ``holes''
in a bottom-up way. For instance, \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{\mid}}\\\;\HOLFreeVar{P}\ensuremath{)}}
holds because \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}}
holds as a base case of the inductive definition of \HOLinline{\HOLConst{WG}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{a}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}\hfill{[WG1]}
\end{alltt}
For any agent variable $X$, we have by $\beta$-reduction:
\HOLinline{\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}\ensuremath{)}\\\;\ensuremath{(}\HOLConst{var}\\\;\HOLFreeVar{X}\ensuremath{)}\\\;\HOLSymConst{\ensuremath{=}}\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}} (or ``$l.X + P$'' in textbook
notation) is weakly guarded (cf.~Def.~\ref{def:guardness}).

@@

Not every expression fit with CCS syntax is a context: any free
variable in concern must not occur inside any recusion operator
(cf.~the footnote at the beginning of Section~\ref{ss:SysEq}). In the
extreme case, if the set of free variables of an expression, say $E$,
is disjoint with a list of variables $\til X$, then this expression is
for sure a context:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{E}\hfill{[disjoint_imp_context]}
\end{alltt}
On the other hand, for any context (w.r.t.~$\til X$) of the form $\recu Y E$, we can
conclude that, the set of free variables of $E$ excluding $Y$, is
disjoint with $\til X$, i.e.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{context} \HOLFreeVar{Xs} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{Y} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLFreeVar{E} \HOLConst{DELETE} \HOLFreeVar{Y}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)}\hfill{[context_rec]}
\end{alltt}
Notice that, the above result is the best possible one. We cannot
conclude \HOLinline{\HOLConst{DISJOINT}\\\;\ensuremath{(}\HOLConst{FV}\\\;\HOLFreeVar{E}\ensuremath{)}\\\;\ensuremath{(}\HOLConst{set}\\\;\HOLFreeVar{Xs}\ensuremath{)}}, because $Y$ as a bound
variable of $\recu Y E$ may be also a free variable in $\til X$. We
also cannot conclude \HOLinline{\HOLConst{context}\\\;\HOLFreeVar{Xs}\\\;\HOLFreeVar{E}}, because in $\recu Y E$
it is possible that the bound variable $Y$ occurs inside another
recursion operator, then in $E$ it now becomes free. (See $\recu Y
(\recu Z (a.Y + b.Z) + c.Y)$ for such an example.)

### 2019-11-26 ic-journal/part2.tex
It is recommended that "cf." be used only to suggest a comparison, and the word "see" be used to point to a source of information.
https://en.wikipedia.org/wiki/List_of_Latin_abbreviations

Then, by a structural induction on weakly-guarded context
(\texttt{WG}), we have proved Milner's strong unique-solution theorem
in the \univariate case (c.f.~Section~\ref{sec:multivariate} for the
\multivariate case). The formal proof basically follows the outline of its
informal version. It is tedious due to large amounts of case analyses
on each CCS operator.

>>>>

Then, by a structural induction on weakly-guarded context
(\texttt{WG}), we have proved Milner's strong unique-solution theorem
in the \univariate case (cf.~Section~\ref{sec:multivariate} for the
\multivariate case). The formal proof basically follows the outline of its
informal version. It is tedious due to large amounts of case analyses
on each CCS operator.
